---
title: "Missing data"
author: "Orhan"
date: "04 03 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

## Introduction

libraries

```{r}
library(tidyverse)
library(DT)
```

```{r}
missing_data <- readr::read_csv("https://raw.githubusercontent.com/VIS-SIG/Wonderful-Wednesdays/master/data/2021/2021-02-10/missing_data.csv")
missing_data %>% datatable()
```

In order to think **clearly**, first, I want to drop "*pain.bin"* variables.

```{r}
data <- missing_data %>%
  select(-starts_with("pain.bin"))
data %>% datatable()
```

better to use long format

```{r}
data_long <- data %>%
  mutate(patient = 1:nrow(data)) %>%
  pivot_longer(
    cols = starts_with("pain."),
    names_to = "visit",
    names_prefix = "pain.",
    values_to = "pain",
    values_drop_na = FALSE
  ) %>%
  mutate(visit = as.integer(visit))
```

Lets look at the data as a heatmap.

LOOK AT NA PATTERNS. DO THEY CONTINUE ONCE OCCURED OR RANDOM?

```{r}
data_long %>% 
  #filter(patient<101) %>% 
  ggplot(aes(x = as.character(visit), y = patient)) +
  geom_tile(aes(fill = pain)) +
  scale_x_discrete(limits = as.character(seq(0, 10, by = 1)))
```

patient_data

```{r}
patient_data <- data_long %>%
  group_by(patient, gender, age, bmi, trt) %>%
  summarise(sum = sum(!is.na(pain))) %>%
  ungroup()
```

average non-NA's by gender

```{r}
patient_data %>%
  group_by(gender) %>%
  summarise(mean = mean(sum))
```

average non-NA's by age

```{r}
patient_data %>%
  group_by(age) %>%
  summarise(mean = mean(sum), patientCount = n()) %>%
  ggplot(aes(x = age)) +
  geom_line(aes(y = mean)) +
  geom_col(aes(y = patientCount), fill = "red", alpha = 0.4)
```

average NA's of below 55 and above 55

```{r}
patient_data %>%
  mutate(naCount = 11 - sum) %>%
  mutate(plus55=ifelse(age>55,"+55","<55")) %>% 
  group_by(plus55) %>%
  summarise(mean = mean(naCount))
```

distribution of NA's by ages

```{r}
patient_data %>%
  mutate(naCount = 11 - sum) %>%
  group_by(age) %>%
  summarise(naCount = sum(naCount)) %>%
  ggplot(aes(x = age)) +
  geom_line(aes(y = naCount))
```

lorenz

```{r}
patient_data %>%
  mutate(naCount = 11 - sum) %>% 
  arrange(age) %>% 
  mutate(cumsum=cumsum(naCount)) %>% 
  ggplot()+geom_line(aes(x=age,y=cumsum))
```

distribution of NA's by ages

```{r}
patient_data %>%
  mutate(naCount = 11 - sum) %>%
  mutate(plus55=ifelse(age>55,"+55","younger")) %>% 
  group_by(plus55) %>%
  summarise(naCount = sum(naCount)) %>%
  mutate(x=1) %>% 
  ggplot(aes(fill=plus55,x=x,y=naCount)) +
  geom_col(position="fill", stat="identity")
```

average non-NA's by bmi

```{r}
patient_data %>%
  group_by(bmi) %>%
  summarise(mean = mean(sum)) %>%
  ggplot(aes(x = bmi, y = mean)) +
  geom_line()
```

average non-NA's by trt

```{r}
patient_data %>%
  group_by(trt) %>%
  summarise(mean = mean(sum))
```

first thoughts;

1.  number of non-NA values for each patient histogram
2.  pain data by visit non-NA histogram
3.  arrange by trt to compare treatments

number of non-NA values for each patient histogram

```{r}
data_long %>%
  group_by(patient) %>%
  summarise(sum = sum(!is.na(pain))) %>%
  group_by(sum) %>%
  summarise(n = n()) %>%
  arrange(sum) %>%
  ggplot(aes(x = sum, y = n)) +
  geom_col() +
  scale_x_discrete(name = "number of non-NA's",limits=c(0,11)) +
  scale_y_discrete(name = "number of patients")+
  coord_flip()
```

number of non-NA values for each patient histogram

```{r}
data_long %>%
  mutate(plus55=ifelse(age>55,"+55","-55")) %>% 
  group_by(patient,plus55) %>%
  summarise(sum = sum(is.na(pain))) %>%
  group_by(sum,plus55) %>%
  summarise(n = n()) %>%
  arrange(sum) %>%
  ggplot(aes(x = sum, y = n)) +
  geom_col() +
  scale_x_discrete(name = "number of non-NA's",limits=c(0,11)) +
  scale_y_discrete(name = "number of patients")+
  coord_flip()
```

number of NA values for each patient histogram by age group

```{r}
data_long %>%
  mutate(plus55=ifelse(age>55,"+55","-55")) %>% 
  group_by(patient,plus55) %>%
  summarise(sum = sum(is.na(pain))) %>%
  group_by(sum,plus55) %>%
  summarise(n = n()) %>% 
  mutate(n=ifelse(plus55=="-55",n*-1,n)) %>% 
  ungroup() %>% 
  complete(sum,plus55) %>% 
  mutate(n=replace_na(n,0)) %>% 
  ggplot(aes(x=sum,y=n))+
    geom_col(aes(group=plus55,fill=plus55)) + 
    scale_fill_brewer(palette="Dark2")+
    scale_x_continuous(breaks = seq(0, 11, 1))+
    scale_y_continuous(breaks = seq(-200, 200, 50), 
                     labels = abs(seq(-200, 200, 50)),
                     limits = c(-200,200))+
    coord_flip()+
    labs(x = "Number of NA's", 
         y = "Number of Patients", 
         title = "Number of Patients by total number of NA's\nfor below 55 and above 55 age groups")+ 
    theme_minimal()+
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none")+
    annotate(geom="text", x=5, y=100, label="+55", colour="black",
             size=15, family="Courier", fontface="bold",alpha=0.5)+
    annotate(geom="text", x=5, y=-100, label="-55", colour="black",
             size=15, family="Courier", fontface="bold",alpha=0.5)
```

number of NA values for each patient histogram by age group (without zero's)

```{r}
data_long %>%
  mutate(plus55=ifelse(age>55,"+55","-55")) %>% 
  group_by(patient,plus55) %>%
  summarise(sum = sum(is.na(pain))) %>%
  filter(sum>0) %>% 
  group_by(sum,plus55) %>%
  summarise(n = n()) %>% 
  mutate(n=ifelse(plus55=="-55",n*-1,n)) %>% 
  ungroup() %>% 
  complete(sum,plus55) %>% 
  mutate(n=replace_na(n,0)) %>% 
  ggplot(aes(x=sum,y=n))+
    coord_flip()+
    geom_col(aes(group=plus55,fill=plus55)) + 
    scale_x_continuous(breaks = seq(0, 11, 1))+
    scale_y_continuous(breaks = seq(-30, 30, 10), 
                     labels = abs(seq(-30, 30, 10)),
                     limits = c(-30,30))
```

pain data by visit non-NA histogram

```{r}
patientCount <- data_long %>%
  group_by(patient) %>%
  summarise(n = n()) %>%
  nrow()

data_long %>%
  group_by(visit) %>%
  summarise(sum = sum(is.na(pain))) %>%
  mutate(rate = sum / patientCount) %>%
  ggplot(aes(x = visit, y = rate)) +
  geom_col() +
  scale_x_discrete(name = "visit order") +
  scale_y_continuous(name = "number of NA's", limits = c(0, 1))
```

arrange by trt to compare treatments

```{r}
data_long %>%
  arrange(trt) %>%
  ggplot(aes(x = as.character(visit), y = patient)) +
  geom_tile(aes(fill = pain)) +
  scale_x_discrete(limits = as.character(seq(0, 10, by = 1)))
```
